Probabilistic Catalogs from Crowded Star Fields

### 2012-10-08 paper/ms.tex
Got rid of some whitespace

\section{Broken Power-Law Distribution}\label{power_law}
The broken power-law distribution is based on a straightforward extension
to a simple power-law distribution (also known as a Pareto distribution,
particularly in the statistics literature). The power-law distribution for a
variable $x$ (given a lower cutoff $x=h$ and a slope $\alpha$) is defined by:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcr}
0, & & x < h \\
x^{-\alpha - 1} & & x \geq h.
\end{array}
\right.
\end{eqnarray}
In contrast, the broken power-law distribution for a variable $x$ is defined by
a lower cutoff $x=h_1$, two slopes $\{\alpha_1, \alpha_2\}$ and a break point
$x=h_2$:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
x^{-\alpha_1 - 1} & & h_1 \leq x \leq h_2 \\
x^{-\alpha_2 - 1} & & x > h_2.
\end{array}
\right.
\end{eqnarray}
The free parameters of the broken power-law are:
\begin{eqnarray}
\beta = \{h_1, h_2, \alpha_1, \alpha_2\}
\end{eqnarray}

\begin{figure}
\begin{center}
\includegraphics[scale=0.7]{Figures/broken.eps}
\end{center}
\end{figure}

With normalising terms included, the proportionality becomes an equality:
\begin{eqnarray}
p(x) &=&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
Z_1^{-1}x^{-\alpha_1 - 1}, & & h_1 \leq x \leq h_2 \\
Z_2^{-1}x^{-\alpha_2 - 1}, & & x > h_2
\end{array}
\right.
.
\end{eqnarray}

Two conditions will be used to determine the normalisers $Z_1$ and $Z_2$.
Firstly, the PDF should be continuous at $x=h_2$:
\begin{eqnarray}
Z_1^{-1}h_2^{-\alpha_1 - 1} &=& Z_2^{-1}h_2^{-\alpha_2 - 1}\\
\implies
Z_2 &=& Z_1h_2^{\alpha_1-\alpha_2}
\end{eqnarray}

The second condition is that the total probability must be 1:

\begin{eqnarray}
\int_{h_1}^{h_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
+
\int_{h_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_2^{-1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_1^{-1}h_2^{\alpha_2-\alpha_1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1
\end{eqnarray}

\begin{eqnarray}
\implies
Z_1 &=& \alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
h_2^{-\alpha_1}\alpha_2^{-1}
\end{eqnarray}

The cumulative distribution (CDF) is:
\begin{eqnarray}
P(X \leq x) = F(x) &=& 
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
(Z_1\alpha_1)^{-1}\left(h_1^{-\alpha_1} - x^{-\alpha_1}\right), & & h_1 \leq x \leq h_2 \\
1 - (Z_2\alpha_2)^{-1}x^{-\alpha_2}, & & x > h_2 
\end{array}
\right.
\end{eqnarray}
and the inverse CDF is:
\begin{eqnarray}
F^{-1}(u) &=& 
\left\{
\begin{array}{lcr}
\left[h_1^{-\alpha_1} - uZ_1\alpha_1\right]^{-1/\alpha_1}, & & 0 < u < 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}\\
\left[Z_2\alpha_2(1-u)\right]^{-1/\alpha_2},& & 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}< u < 1
\end{array}
\right.
\end{eqnarray}

>>>>

\section{Broken Power-Law Distribution}\label{power_law}
The broken power-law distribution is based on a straightforward extension
to a simple power-law distribution (also known as a Pareto distribution,
particularly in the statistics literature). The power-law distribution for a
variable $x$ (given a lower cutoff $x=h$ and a slope $\alpha$) is defined by:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcr}
0, & & x < h \\
x^{-\alpha - 1} & & x \geq h.
\end{array}
\right.
\end{eqnarray}
In contrast, the broken power-law distribution for a variable $x$ is defined by
a lower cutoff $x=h_1$, two slopes $\{\alpha_1, \alpha_2\}$ and a break point
$x=h_2$:
\begin{eqnarray}
p(x) &\propto&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
x^{-\alpha_1 - 1} & & h_1 \leq x \leq h_2 \\
x^{-\alpha_2 - 1} & & x > h_2.
\end{array}
\right.
\end{eqnarray}
The free parameters of the broken power-law are:
\begin{eqnarray}
\beta = \{h_1, h_2, \alpha_1, \alpha_2\}.
\end{eqnarray}
With normalising terms included, the proportionality becomes an equality:
\begin{eqnarray}
p(x) &=&
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
Z_1^{-1}x^{-\alpha_1 - 1}, & & h_1 \leq x \leq h_2 \\
Z_2^{-1}x^{-\alpha_2 - 1}, & & x > h_2.
\end{array}
\right.
\end{eqnarray}
Two conditions will be used to determine the normalisers $Z_1$ and $Z_2$.
Firstly, the PDF should be continuous at $x=h_2$:
\begin{eqnarray}
Z_1^{-1}h_2^{-\alpha_1 - 1} &=& Z_2^{-1}h_2^{-\alpha_2 - 1}\\
\implies
Z_2 &=& Z_1h_2^{\alpha_1-\alpha_2}
\end{eqnarray}
The second condition is that the total probability must be 1:
\begin{eqnarray}
\int_{h_1}^{h_2} Z_1^{-1} x^{-\alpha_1 - 1} \, dx
+
\int_{h_2}^\infty Z_2^{-1} x^{-\alpha_2 - 1} \, dx
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_2^{-1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1 \\
Z_1^{-1}\alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
Z_1^{-1}h_2^{\alpha_2-\alpha_1}\alpha_2^{-1}h_2^{-\alpha_2}
&=& 1
\end{eqnarray}
\begin{eqnarray}
\implies
Z_1 &=& \alpha_1^{-1}\left[h_1^{-\alpha_1} - h_2^{-\alpha_1}\right]
+
h_2^{-\alpha_1}\alpha_2^{-1}.
\end{eqnarray}
The cumulative distribution (CDF) is:
\begin{eqnarray}
P(X \leq x) = F(x) &=& 
\left\{
\begin{array}{lcr}
0, & & x < h_1 \\
(Z_1\alpha_1)^{-1}\left(h_1^{-\alpha_1} - x^{-\alpha_1}\right), & & h_1 \leq x \leq h_2 \\
1 - (Z_2\alpha_2)^{-1}x^{-\alpha_2}, & & x > h_2 
\end{array}
\right.
\end{eqnarray}
and the inverse CDF is:
\begin{eqnarray}
F^{-1}(u) &=& 
\left\{
\begin{array}{lcr}
\left[h_1^{-\alpha_1} - uZ_1\alpha_1\right]^{-1/\alpha_1}, & & 0 < u < 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}\\
\left[Z_2\alpha_2(1-u)\right]^{-1/\alpha_2},& & 1 - (Z_2\alpha_2)^{-1}h_2^{-\alpha_2}< u < 1
\end{array}
\right.
\end{eqnarray}
An example of a broken power-law distribution is shown in Figure~\ref{fig:powerlaw}.
\begin{figure}[h!]
\begin{center}
\includegraphics[scale=0.7]{Figures/broken.eps}
\caption{A broken power-law distribution.\label{fig:powerlaw}}
\end{center}
\end{figure}
